---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# OLStrajr

<!-- badges: start -->
<!-- badges: end -->

The [Maximizing the Yield of Small Samples in Prevention Research: A Review of General Strategies and Best Practices](https://doi.org/10.1007%2Fs11121-014-0542-7) paper by Hopkin, Hoyle, and Gottfredson(2015) suggests that when working with small sample sizes, analysts should leverage plots to explore within-group associations between predictor variables and the outcome variable. These plots are intended to provide detailed insights about the data rather than to be used for generalizing to the population. 

The paper cites the work of [Carrig, Wirth, and Curran (2004)](http://dx.doi.org/10.1207/S15328007SEM1101_9) who proposed an easy-to-use SAS macro for visualizing person-specific growth trajectories with repeated measures data. 

The objective of the OLStrajr package is to bring the functionality of this SAS macro, OLStraj, to R. While the original link to the macro in Carrig's paper is no longer active, it can still be accessed through the [Wayback Machine](https://archive.org/web/).

## Installation

You can install the development version of OLStrajr as follows:

``` {r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("mightymetrika/OLStrajr")
```

## GPT-4 Dog Barking Poisson

This is an illustrative example that shows how to use OLStrajr.

On June 12th, 2023, I asked GPT-4 with Browsing to provide an estimation of the number of times a dog might bark from 6:00AM to 6:01AM. I received the following response:

"The number of times a dog barks within a minute can vary widely based on many factors, such as the dog's breed, personality, and the stimulus causing the barking. Some dogs may bark once or twice and then stop, while others may bark continuously. If we take a rough estimate, a dog that barks continuously might bark once every two seconds, which would equate to about 30 barks in a minute. However, this is a very rough estimate and the actual number could be much higher or lower depending on the dog and the circumstances.

Please note that excessive barking can be a sign of stress, boredom, or other issues in a dog, so if a dog is barking a lot, it's important to try to identify and address the underlying cause."

Based on this response, I simulated five days of data from five different houses, where the dogs start barking at 6AM. The barking frequency is modeled as a Poisson process where the week becomes increasingly boring. This basic example shows how to use OLStrajr

```{r}
library(OLStrajr)
```


### Simulate Data

```{r}
set.seed(489)
barks <- data.frame(dog = c(1, 2, 3, 4, 5),
                    mon = rpois(5, 14),
                    tues = rpois(5, 18),
                    weds = rpois(5, 22),
                    thurs = rpois(5, 26),
                    fri = rpois(5, 30))
```


### OLStraj

To obtain the OLS trajectories, use the following code:

```{r}
barks_traj <- OLStraj(data = barks,
                      idvarname = "dog",
                      predvarname = "Day",
                      outvarname = "Barks",
                      varlist = c("mon", "tues", "weds", "thurs", "fri"),
                      timepts = c(1, 2, 3, 4, 5),
                      regtype = "lin",
                      int_bins = 5,
                      lin_bins = 5)
```


The OLStraj function provides several types of plots that facilitate the examination of your data:

### Group Plots:

```{r}
barks_traj$group_plots
```


Group plots depict the simple-joined trajectories and the estimated trajectories for each group in your data, allowing for a quick comparison of trends across different groups.

### Individual OLS Trajectories:

```{r}
barks_traj$individual_plots
```


Individual plots represent each subject's unique trajectory over time, providing insights into individual patterns of change.

### Histograms for intercepts and slopes:

```{r}
barks_traj$histogram_plots
```


These histograms reveal the distribution of intercepts and slopes across subjects. They are useful for identifying skewness or other interesting distributional properties in the data.

### Box plots of intercepts and slopes:

```{r}
barks_traj$box_plot
```


Box plots provide a summary of the distribution of intercepts and slopes across subjects. They allow you to visualize the spread and skewness of the data. Any outliers are clearly highlighted in these plots. In the OLStraj box plots, the mean value is represented by a blue dot. Additionally, each outlier is labeled with the corresponding subject identifier, providing a straightforward way to identify subjects whose data deviate significantly from the rest of the group.
